---
title: "Calf Health and IgG Analysis"
author: "WAK"
date: "today"
format: 
  html:
    toc: true 
    toc-float: true
    toc-location: left
---

## Data Steps

```{r}
#| include: false
#| warning: false
#| echo: false

library(epiDisplay)
library(knitr)
library(pROC)
library(tidyverse)
library(readxl)
library(ggplot2)
library(gmodels)
library(ggbeeswarm)
library(dplyr)
library(ggpubr)
library(psych)
library(lme4)
library(Matrix)
library(lmerTest)
library(irr)
library(lmerTest)
library(emmeans)
library(doBy)       
library(gmodels)    
library(car)           
library(cowplot)       
library(gridGraphics)  
library(multcomp)
library(hms)
library(gt)
library(RColorBrewer)
library(extrafont)
library(nlme)
```

```{r}
#| include: false
#| echo: false
#| warning: false

enroll <- read_csv('data/Zinpro 2024 enrollment final.csv')
rmm<- read_csv('data/RMM.csv')


clean <-
  enroll |> 
  filter(CalfID!=113)

rmm_clean <-
  rmm |> 
  filter(CalfID != 113)

clean<- clean %>% 
  mutate(
    bwkg = BirthWtLbs/2.2, 
    percFirst = Fed1VolL/bwkg, 
    D0Fed2VolConsumL = DOFed2VolConsummL/1000,
    DOFed2VolOfferedL = DOFed2VolOfferedmL/1000,
    percSecond = D0Fed2VolConsumL/bwkg, 
    totaloffered = Fed1VolL + DOFed2VolOfferedL)

#removes calf 113, who did not have correct information    

#changes these variables to "hours" vs. "seconds"
clean$DOFed2HrsOld <- as.numeric(clean$DOFed2HrsOld)/3600
clean$DOFed12IntervHrs <- as.numeric(clean$DOFed12IntervHrs)/3600
clean$DOFed1MinOld <- as.numeric(clean$DOFed1MinOld)/60
  
clean<- 
  clean |> 
  mutate(refuse=
           case_when(
             D0Fed2VolRefusemL > 0    ~ 1, 
             D0Fed2VolRefusemL == 0   ~ 0))

clean$refuse <- as.factor(clean$refuse)

clean<- 
  clean |> 
  mutate(drankall=
           case_when(
             D0Fed2VolRefusemL > 0    ~ 0, 
             D0Fed2VolRefusemL == 0   ~ 1))

clean$drankall <- as.factor(clean$drankall)


clean<- 
  clean |> 
  mutate(refuseM1=
           case_when(
             D1MilkRefusemL > 0    ~ 1, 
             D1MilkRefusemL == 0   ~ 0))

clean<- 
  clean |> 
  mutate(
    propconsume = DOFed2VolConsummL/DOFed2VolOfferedmL
  )

clean<-
  clean |> 
  mutate(TotalVol = (D0Fed13or4qt*946.35) + DOFed2VolConsummL)



IgG <- 
  clean |> 
  select ( D0Fed2TxGrp, Result8hrgL, Result24hrgL) |> 
  pivot_longer(
  cols = Result8hrgL:Result24hrgL) 


MCCS <- 
  clean |> 
  mutate(
    type = case_when(
      D0Fed2TxGrp == "MCB"  ~ "COL",
      D0Fed2TxGrp == "MCT"  ~ "COL", 
      D0Fed2TxGrp == "CST"  ~ "CS", 
      D0Fed2TxGrp == "CSB"  ~ "CS", 
      D0Fed2TxGrp == "MRB"  ~ "NONE"))


clean <- clean |> 
  mutate(
    botvstube = case_when(
      D0Fed2TxGrp == "MCB"  ~ "BOT",
      D0Fed2TxGrp == "MCT"  ~ "TUB", 
      D0Fed2TxGrp == "CST"  ~ "TUB", 
      D0Fed2TxGrp == "CSB"  ~ "BOT", 
      D0Fed2TxGrp == "MRB"  ~ "BOT"))

```

## Descriptive Analysis

### Calf Enrollment Characteristics

This generally describes calves in each treatment group by various indicators (breed, sex, size). Calves were balanced between treatment groups (n=31). Majority of enrolled calves were crossbreds, and fairly evenly split between large and small, and male and female. There was one calf who did not have their sex recorded.

```{r}
#| echo: false
#| warning: false

n_group <- 
  clean |> 
  group_by(D0Fed2TxGrp) |> 
  summarize (Count = n())
 
n_group

summary_sex <- 
  clean |> 
  group_by(D0Fed2TxGrp, SexMF) |> 
  summarize(Count = n())

table_sex <- pivot_wider(summary_sex, names_from = SexMF, values_from = Count, values_fill = 0)
table_matrix <- as.matrix(table_sex[, -1])  
chi_result <- chisq.test(table_matrix)
chi_result


summary_cb <- 
  clean |> 
  group_by(D0Fed2TxGrp, CrossbredYN) |> 
  summarize(Count = n()) 

table_cb <- pivot_wider(summary_cb, names_from = CrossbredYN, values_from = Count, values_fill = 0)
table_matrix <- as.matrix(table_cb[, -1])  
chi_result <- chisq.test(table_matrix)
chi_result


summary_size<-
  clean |> 
  group_by(D0Fed2TxGrp, CalfSizeLgSm) |> 
  summarize(Count=n()) 

table_size <- pivot_wider(summary_size, names_from = CalfSizeLgSm, values_from = Count, values_fill = 0)
table_matrix <- as.matrix(table_size[, -1])  
chi_result <- chisq.test(table_matrix)
chi_result


summary_twin<-
  clean |> 
  group_by(D0Fed2TxGrp, SingleTwinST) |> 
  summarize(Count=n()) 

table_twin <- pivot_wider(summary_twin, names_from = SingleTwinST, values_from = Count, values_fill = 0)
table_matrix <- as.matrix(table_twin[, -1])  
chi_result <- chisq.test(table_matrix)
chi_result


summary_replace <- 
  clean |> 
  group_by(D0Fed2TxGrp, Replacement) |> 
  summarize(Count = n())

table_replace <- pivot_wider(summary_replace, names_from = Replacement, values_from = Count, values_fill = 0)
table_matrix <- as.matrix(table_replace[, -1])  
chi_result <- chisq.test(table_matrix)
chi_result




sick_<-
  clean |> 
  group_by(D0Fed2TxGrp, FarmDzYN) |> 
  summarize(count=n(), .groups="drop") |> 
  mutate(proportion = signif(count/sum(count), digits=2)) 
  



n_group
summary_sex
summary_cb
summary_size
summary_twin
summary_replace
sick_
```

### Birth Weight by Treatment Group

Calf birth weight was relatively consistent between treatment groups. Calves weighed between 83 and 87lbs at birth/enrollment.

```{r}
#| echo: false
#| warning: false


clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(BirthWtLbs)/2.2, 
             sd = sd(BirthWtLbs)/2.2)

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_bw <- lm(BirthWtLbs/2.2 ~ D0Fed2TxGrp, data=clean)
summary(lm_bw)
  

clean |> 
  group_by (D0Fed2TxGrp, CalfSizeLgSm) |> 
  summarize (mean = mean(BirthWtLbs), 
             sd = sd(BirthWtLbs))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_bwsize <- lm(BirthWtLbs ~ D0Fed2TxGrp*CalfSizeLgSm, data=clean)
summary(lm_bwsize)



clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = BirthWtLbs, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter(alpha = 0.75) +
  scale_color_manual(
    values = c(
      "MRB" = "grey30",
      "CSB" = "gold",
      "CST" = "gold3",
      "MCB" = "red3",
      "MCT" = "red4"))+ 
  labs(x = "Second Feeding Treatment", 
       y = "Birth Weight (lbs)")+
  scale_y_continuous(limits = c(0, 150), breaks = seq(0, 150, by = 25)) +
  theme(legend.position = "none", 
        text = element_text(family = "Times New Roman", size = 12))
```

### Serum IgG at 0 hrs (g/L)

```{r}
clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(Result0hrgL), 
             sd = sd(Result0hrgL))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_IgG0 <- lm(Result0hrgL ~ D0Fed2TxGrp, data=clean)
summary(lm_IgG0)
```

## FIRST FEEDING

### Age at First Feeding(min)

```{r}
#| echo: false
#| warning: false

clean |> 
  summarize(mean=mean(DOFed1MinOld), 
            sd = sd (DOFed1MinOld)) 


clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(DOFed1MinOld), 
             sd = sd(DOFed1MinOld))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_minold <- lm(DOFed1MinOld ~ D0Fed2TxGrp, data=clean)
summary(lm_minold)

clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = DOFed1MinOld, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
   scale_color_manual(
    values = c(
      "MRB" = "grey30",
      "CSB" = "gold",
      "CST" = "gold3",
      "MCB" = "red3",
      "MCT" = "red4"))+ 
  labs(x = "Second Feeding Treatment", 
       y = "Age at first feeding(min)") +
  theme(legend.position = "none", 
        text = element_text(family = "Times New Roman", size = 12))

```

### Colostral IgG Feeding 1

```{r}
#| echo: false
#| warning: false

clean |> 
  summarize (mean = mean(F1ColostIgGgL), 
             sd = sd(F1ColostIgGgL))

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(F1ColostIgGgL), 
             sd = sd(F1ColostIgGgL))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_F1Colost <- lm(F1ColostIgGgL ~ D0Fed2TxGrp, data=clean)
summary(lm_F1Colost)

clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = F1ColostIgGgL, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
   scale_color_manual(
    values = c(
      "MRB" = "grey30",
      "CSB" = "gold",
      "CST" = "gold3",
      "MCB" = "red3",
      "MCT" = "red4"))+ 
  labs(x = "Second Feeding Treatment", 
       y = "First Feeding IgG in Colostrum")+
  scale_y_continuous(limits = c(0, 125), breaks = seq(0, 125, by = 25))+
   theme(legend.position = "none", 
        text = element_text(family = "Times New Roman", size = 12))

```

### Feeding 1 Total IgG Dose

```{r}
#| echo: false
#| warning: false

clean |> 
  summarize (mean = mean(F1DoseIgGfedg), 
             sd = sd(F1DoseIgGfedg))

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(F1DoseIgGfedg), 
             sd = sd(F1DoseIgGfedg))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_F1Dose <- lm(F1DoseIgGfedg ~ D0Fed2TxGrp, data=clean)
summary(lm_F1Dose)

clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = F1DoseIgGfedg, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
     scale_color_manual(
    values = c(
      "MRB" = "grey30",
      "CSB" = "gold",
      "CST" = "gold3",
      "MCB" = "red3",
      "MCT" = "red4"))+ 
  labs(y = "First Feeding IgG Dose",
       x = NULL)+
  scale_y_continuous(limits = c(0, 450), breaks = seq(0, 450, by = 50))+ theme(legend.position = "none", 
        text = element_text(family = "Times New Roman", size = 12))
```

### Fed 1 BBW%

```{r}
#| echo: false
#| warning: false

clean |> 
  summarize (mean = mean(percFirst), 
             sd = sd(percFirst))


clean |> 
  group_by(CalfSizeLgSm) %>% 
  summarize (mean = mean(percFirst),
             sd = sd(percFirst), 
             meanv = mean(Fed1VolL), 
             sdv = sd(Fed1VolL))




clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(percFirst), 
             sd = sd(percFirst))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_F1perc <- lm(percFirst ~ D0Fed2TxGrp, data=clean)
summary(lm_F1perc)
```

### Assn between bacteria and Colostral IgG

```{r}
#| echo: false
#| warning: false

lm_coliform <- lm(F1ColostIgGgL~F1LogColif, data=clean)
summary(lm_coliform)

clean |> 
ggplot(mapping = aes(x = F1ColostIgGgL, y = F1LogColif)) +
  theme_classic()+
  geom_point() +
  geom_smooth(method = "lm", se=TRUE, color="red")+
  labs(x = "Colostral IgG", y = "Log Coliform Count", title = "Scatterplot of Colostral IgG vs. Coliform")




lm_spc <- lm(F1ColostIgGgL~F1SPC, data=clean)
summary(lm_spc)

clean |> 
ggplot(mapping = aes(x = F1ColostIgGgL, y = F1SPC)) +
  theme_classic()+
  geom_point() +
  geom_smooth(method = "lm", se=TRUE, color="red")+
  labs(x = "Colostral IgG", y = "SPC", title = "Scatterplot of Colostral IgG vs. SPC")
```

### Brix% vs Colostral IgG

```{r}
#| echo: false
#| warning: false

lm_brix <- lm(F1ColostIgGgL~F1Brixpct, data=clean)
summary(lm_brix)

clean |> 
ggplot(mapping = aes(x = F1ColostIgGgL, y = F1Brixpct)) +
  theme_classic()+
  geom_point() +
  geom_smooth(method = "lm", se=TRUE, color="red")+
  labs(x = "Colostral IgG", y = "Brix %", title = "Scatterplot of Colostral IgG vs. Brix")
```

## SECOND FEEDING

### Hrs Old At Second Feeding

```{r}
#| echo: false
#| warning: false

clean |> 
  summarize (mean = mean(DOFed2HrsOld), 
             sd = sd(DOFed2HrsOld))

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(DOFed2HrsOld), 
             sd = sd(DOFed2HrsOld))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_hrsold <- lm(DOFed2HrsOld ~ D0Fed2TxGrp, data=clean)
summary(lm_hrsold)

clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = DOFed2HrsOld, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  scale_color_manual(
    values = c(
      "MRB" = "grey30",
      "CSB" = "gold",
      "CST" = "gold3",
      "MCB" = "red3",
      "MCT" = "red4"))+ 
  labs(x = NULL, 
       y = "Hours old at second Feeding")+
  theme(legend.position = "none", 
        text = element_text(family = "Times New Roman", size = 12))
  

```

### Interval Feeding 1 - 2

```{r}
#| echo: false
#| warning: false

clean |> 
  summarize (mean = mean(DOFed12IntervHrs), 
             sd = sd(DOFed12IntervHrs))

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(DOFed12IntervHrs), 
             sd = sd(DOFed12IntervHrs))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_intvhrs <- lm(DOFed12IntervHrs ~ D0Fed2TxGrp, data=clean)
summary(lm_intvhrs)

clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = DOFed12IntervHrs, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  scale_color_manual(
    values = c(
      "MRB" = "grey30",
      "CSB" = "gold",
      "CST" = "gold3",
      "MCB" = "red3",
      "MCT" = "red4"))+ 
  labs(x = NULL, 
       y = "Feeding Interval")+
  theme(legend.position = "none", 
        text = element_text(family = "Times New Roman", size = 12))
```

### 8hr IgG by Treatment Group

```{r}
#| echo: false
#| warning: false

clean |> 
  summarize(mean=mean(Result8hrgL, na.rm=TRUE), 
            sd = sd(Result8hrgL, na.rm = TRUE))

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(Result8hrgL, na.rm = TRUE), 
             sd = sd(Result8hrgL, na.rm = TRUE))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_8hr <- lm(Result8hrgL ~ D0Fed2TxGrp, data=clean)
summary(lm_8hr)

clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = Result8hrgL, fill=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot() +
  geom_jitter(alpha = 0.5) +
  scale_fill_manual(
    values = c(
      "MRB" = "grey30",
      "CSB" = "gold",
      "CST" = "gold3",
      "MCB" = "red3",
      "MCT" = "red4")) +
  labs(x = NULL, 
       y = "Serum IgG at 8 hrs")+
  scale_y_continuous(limits = c(0, 60), breaks = seq(0, 60, by = 10)) +
  theme(legend.position = "none", 
        text = element_text(family = "Arial", size = 14))



```

### Refusals, second feeding (mLs and Refusal Y/N)

```{r}
#| echo: false
#| warning: false

clean |> 
  summarize (mean = mean(D0Fed2VolRefusemL), 
             median = median (D0Fed2VolRefusemL),
             min = min(D0Fed2VolRefusemL),
             max = max(D0Fed2VolRefusemL),
             sd = sd(D0Fed2VolRefusemL))

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(D0Fed2VolRefusemL), 
             median = median (D0Fed2VolRefusemL),
             min = min(D0Fed2VolRefusemL),
             max = max(D0Fed2VolRefusemL),
             sd = sd(D0Fed2VolRefusemL))

clean |> 
  group_by (D0Fed2TxGrp) |> 
  count(refuse)

clean |> 
  group_by (D0Fed2TxGrp) |> 
  count(drankall)

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_refuse <- lm(D0Fed2VolRefusemL ~ D0Fed2TxGrp, data=clean)
summary(lm_refuse)


clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = D0Fed2VolRefusemL, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  scale_color_manual(
    values = c(
      "MRB" = "grey30",
      "CSB" = "gold",
      "CST" = "gold3",
      "MCB" = "red3",
      "MCT" = "red4")) +
  labs(x = NULL, 
       y = "Refusals (mLs)")+
  theme(legend.position = "none", 
        text = element_text(family = "Times New Roman", size = 12))


```

### Amount Consumed Second Feeding

```{r}
#| echo: false
#| warning: false

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(DOFed2VolConsummL), 
             sd = sd(DOFed2VolConsummL))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_refuse <- lm(DOFed2VolConsummL ~ D0Fed2TxGrp, data=clean)
summary(lm_refuse)
emmeans(lm_refuse, pairwise ~ D0Fed2TxGrp, adjust = "tukey")



clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = DOFed2VolConsummL, fill=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  scale_fill_manual(
    values = c(
      "MRB" = "grey30",
      "CSB" = "gold",
      "CST" = "gold3",
      "MCB" = "red3",
      "MCT" = "red4")) +
  labs(x = NULL, 
       y = "Amount consumed (mL)")+
   theme(legend.position = "none", 
        text = element_text(family = "Arial", size = 14))


#proportion consumed
clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(propconsume), 
             sd = sd(propconsume))


clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_percon <- lm(propconsume ~ D0Fed2TxGrp, data=clean)
summary(lm_percon)
emmeans(lm_percon, pairwise ~ D0Fed2TxGrp, adjust = "tukey")



clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = propconsume, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  labs(x = "Second Feeding Treatment", 
       y = "Proportion Consumed", 
       title = "Treatment vs. Proportion Consumed")


clean |> 
  group_by (botvstube) |> 
  summarize (mean = mean(propconsume), 
             sd = sd(propconsume))

```

### Fed 2 BBW%

```{r}
#| echo: false
#| warning: false

clean |> 
  summarize (mean = mean(percSecond), 
             sd = sd(percSecond))

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(percSecond), 
             sd = sd(percSecond))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_F2perc <- lm(percSecond ~ D0Fed2TxGrp, data=clean)
summary(lm_F2perc)
emmeans(lm_F2perc, pairwise ~ D0Fed2TxGrp, adjust = "tukey")
```

### Second Feeding Colostral IgG

```{r}
#| echo: false
#| warning: false

MCCS |> 
  group_by(type) |>  
  summarize (mean = mean(F2ColostIgGgL, na.rm=TRUE), 
             sd = sd(F2ColostIgGgL, na.rm = TRUE))

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(F2ColostIgGgL), 
             sd = sd(F2ColostIgGgL))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_colst2 <- lm(F2ColostIgGgL ~ D0Fed2TxGrp, data=clean)
summary(lm_colst2)
emmeans(lm_colst2, pairwise ~ D0Fed2TxGrp, adjust = "tukey")


clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = F2ColostIgGgL, fill=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  scale_fill_manual(
    values = c(
      "MRB" = "grey30",
      "CSB" = "gold",
      "CST" = "gold3",
      "MCB" = "red3",
      "MCT" = "red4")) +
  labs(x = NULL, 
       y = "Second Feeding IgG (g/L)")+
   theme(legend.position = "none", 
        text = element_text(family = "Times New Roman", size = 12))
```

### Second Feeding Dose IgG

Dose this account for calves that didn't drink everything? That is why there is the weird variation?

```{r}
#| echo: false
#| warning: false

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(F2DoseIgGfedg), 
             sd = sd(F2DoseIgGfedg))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_colst2dose <- lm(F2DoseIgGfedg ~ D0Fed2TxGrp, data=clean)
summary(lm_colst2dose)
emmeans(lm_colst2dose, pairwise ~ D0Fed2TxGrp, adjust = "tukey")


clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = F2DoseIgGfedg, fill=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  scale_fill_manual(
    values = c(
      "MRB" = "grey30",
      "CSB" = "gold",
      "CST" = "gold3",
      "MCB" = "red3",
      "MCT" = "red4")) +
  labs(x = NULL, 
       y = "IgG Dose Cosumed (g)")+
   scale_y_continuous(limits = c(0, 150), breaks = seq(0, 150, by = 50)) +
  theme(legend.position = "none", 
        text = element_text(family = "Arial", size = 14))
  

```

## BOTH FEEDINGS

### Time from Birth to Second Feeding, First feeding to Second Feeding

```{r}
#| echo: false
#| warning: false



clean |> 
  summarize(
    meanfull = mean(HrsBtoMilk), 
    sdfull = sd (HrsBtoMilk), 
    meanpart = mean(HrsF2toMilk), 
    sdpart = sd(HrsF2toMilk)
  )


#birth to milk replacer feeding
clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(HrsBtoMilk), 
             sd = sd(HrsBtoMilk))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_full <- lm(HrsBtoMilk ~ D0Fed2TxGrp, data=clean)
summary(lm_full)
emmeans(lm_full, pairwise ~ D0Fed2TxGrp, adjust = "tukey")

#first feeding to milk replacer feeding
clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(HrsF2toMilk), 
             sd = sd(HrsF2toMilk))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_part <- lm(HrsF2toMilk ~ D0Fed2TxGrp, data=clean)
summary(lm_part)
emmeans(lm_part, pairwise ~ D0Fed2TxGrp, adjust = "tukey")
```

### Total Dose IgG Consumed (F1 + F2)

```{r}
#| echo: false
#| warning: false

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(F1F2TotIgGconsumg), 
             sd = sd(F1F2TotIgGconsumg))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_colstdose <- lm(F1F2TotIgGconsumg ~ D0Fed2TxGrp, data=clean)
summary(lm_colstdose)
emmeans(lm_colstdose, pairwise ~ D0Fed2TxGrp, adjust = "tukey")


clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = F1F2TotIgGconsumg, fill=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot() +
  geom_jitter(alpha = 0.5) +
  scale_fill_manual(
    values = c(
      "MRB" = "grey30",
      "CSB" = "gold",
      "CST" = "gold3",
      "MCB" = "red3",
      "MCT" = "red4")) +
  labs(x = NULL, 
       y = "Total IgG Dose (g)")+
  scale_y_continuous(limits = c(0, 600), breaks = seq(0, 600, by = 100))+
  theme(legend.position = "none", 
        text = element_text(family = "Arial", size = 14))
```

### Total Volume Consume (F1+F2)

```{r}
#| echo: false
#| warning: false

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(TotalVol), 
             sd = sd(TotalVol))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_Tvol <- lm(TotalVol ~ D0Fed2TxGrp, data=clean)
summary(lm_Tvol)
emmeans(lm_Tvol, pairwise ~ D0Fed2TxGrp, adjust = "tukey")


clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = TotalVol, fill=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter(alpha = 0.5) +
    scale_fill_manual(
    values = c(
      "MRB" = "grey30",
      "CSB" = "gold",
      "CST" = "gold3",
      "MCB" = "red3",
      "MCT" = "red4")) +
  labs(x = NULL, 
       y = "Total Volume Consumed (mL)")+
  scale_y_continuous(limits = c(0, 7000), breaks = seq(0, 7000, by = 1000))+
  theme(legend.position = "none", 
        text = element_text(family = "Arial", size = 14))
```

### Total % consume

```{r}
#| echo: false
#| warning: false

clean |> 
  summarize (mean = mean(TotalPropConvOffered), 
             sd = sd(TotalPropConvOffered))

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(TotalPropConvOffered), 
             sd = sd(TotalPropConvOffered))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_totalperc <- lm(TotalPropConvOffered ~ D0Fed2TxGrp, data=clean)
summary(lm_totalperc)
emmeans(lm_totalperc, pairwise ~ D0Fed2TxGrp, adjust = "tukey")
```

### 24hr serum IgG

```{r}
#| echo: false
#| warning: false


clean |> 
  summarize (mean = mean(Result24hrgL, na.rm=TRUE), 
             sd = sd(Result24hrgL, na.rm=TRUE))


#one calf in the CSB group that doesn't have a 24hr blood IgG
clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(Result24hrgL, na.rm=TRUE), 
             sd = sd(Result24hrgL, na.rm=TRUE))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_24hr <- lm(Result24hrgL ~ D0Fed2TxGrp, data=clean)
summary(lm_24hr)


clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = Result24hrgL, fill=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter(alpha = 0.5) +
  scale_fill_manual(
    values = c(
      "MRB" = "grey30",
      "CSB" = "gold",
      "CST" = "gold3",
      "MCB" = "red3",
      "MCT" = "red4")) +
  labs(x = NULL, 
       y = "Serum IgG at 24 hrs")+
  scale_y_continuous(limits = c(0, 60), breaks = seq(0, 60, by = 10)) +
  theme(legend.position = "none", 
        text = element_text(family = "Arial", size = 14))
 
#serum IgG if calves had refusals at second feeding

clean |> 
ggplot(mapping = aes(x = refuse, y = Result24hrgL, color=refuse)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  labs(x = "Refusals at second feeding (Y/N)", 
       y = "Serum IgG at 24 hrs", 
       title = "Refusals vs. IgG at 24 hours")




```

## POST COLOSTRUM

### First Milk (only) Feeding - Refusals

```{r}
#| echo: false
#| warning: false

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(D1MilkRefusemL),
             median = median (D1MilkRefusemL),
             sd = sd(D1MilkRefusemL), 
             meanh = mean(HrsF2toMilk))

clean |> 
  summarize (
    meanh = mean(HrsF2toMilk), 
    sd = sd (HrsF2toMilk), 
    min = min(HrsF2toMilk), 
    max = max (HrsF2toMilk))

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (refuse = sum(refuseM1))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_refF2 <- lm(D1MilkRefusemL ~ D0Fed2TxGrp, data=clean)
summary(lm_refF2)


clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = D1MilkRefusemL, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.shape=NA) +
  geom_jitter() +
   scale_color_manual(
    values = c(
      "MRB" = "grey30",
      "CSB" = "gold",
      "CST" = "gold3",
      "MCB" = "red3",
      "MCT" = "red4")) +
  labs(x = NULL, 
       y = "24hr milk refusal (mLs)") +
  theme(legend.position = "none", 
        text = element_text(family = "Times New Roman", size = 12))
```

## MODELS

### IgG at 8Hrs

```{r}
clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_8hr <- lm(Result8hrgL ~ D0Fed2TxGrp, data=clean)
summary(lm_8hr) #no

lm_8hrBreed <- lm(Result8hrgL ~ CrossbredYN, data=clean)
summary(lm_8hrBreed)#no

lm_8hrSize <- lm(Result8hrgL ~ CalfSizeLgSm, data=clean)
summary(lm_8hrSize) #no

lm_8hrBW <- lm(Result8hrgL ~ BirthWtLbs, data=clean)
summary(lm_8hrBW)#no


lm_8hrSex <- lm(Result8hrgL ~ SexMF, data=clean)
summary(lm_8hrSex) #yes

lm_8hrST <- lm(Result8hrgL ~ SingleTwinST, data=clean)
summary(lm_8hrST)#no

lm_minold8 <- lm(Result8hrgL ~ DOFed1MinOld, data=clean)
summary(lm_minold8) #no

lm_perc <- lm(Result8hrgL ~ percFirst, data=clean)
summary(lm_perc)#not associated

lm_dose <- lm(Result8hrgL ~ F1DoseIgGfedg, data=clean)
summary(lm_dose)#not associated

lm_rep <- lm(Result8hrgL ~ Replacement, data=clean)
summary(lm_rep)#nope



#model
glmIgG8 <- lm(Result8hrgL ~ D0Fed2TxGrp + SexMF, data=clean)
summary(glmIgG8)
plot(glmIgG8)
car::Anova(glmIgG8, type=3)


emmeans (glmIgG8, ~ D0Fed2TxGrp)
emmeans(glmIgG8, ~SexMF)
x<- emmeans (glmIgG8, ~ D0Fed2TxGrp)
pairwise_comparisons2 <- pairs(x)
summary(pairwise_comparisons2, adjust="tukey")
summary(pairwise_comparisons2)

```

### IgG at 24Hrs

```{r}

#Associations between IgG at 24Hr and Calf Demographics



clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_24hr <- lm(Result24hrgL ~ D0Fed2TxGrp, data=clean)
summary(lm_24hr) #no

lm_24hrBreed <- lm(Result24hrgL ~ CrossbredYN, data=clean)
summary(lm_24hrBreed)#yes

lm_24hrSize <- lm(Result24hrgL ~ CalfSizeLgSm, data=clean)
summary(lm_24hrSize) #no

lm_24hrBW <- lm(Result24hrgL ~ BirthWtLbs, data=clean)
summary(lm_24hrBW)#no

lm_24hrRefuse <- lm(Result24hrgL ~ refuse, data=clean)
summary(lm_24hrRefuse)#yes

lm_24hrSex <- lm(Result24hrgL ~ SexMF, data=clean)
summary(lm_24hrSex) #yes

lm_24hrST <- lm(Result24hrgL ~ SingleTwinST, data=clean)
summary(lm_24hrST)#no

lm_minold <- lm(Result24hrgL ~ DOFed1MinOld, data=clean)
summary(lm_minold) #no

lm_8hrv24<- lm(Result24hrgL ~ Result8hrgL, data=clean)
summary(lm_8hrv24) #no 

lm_perc <- lm(Result24hrgL ~ percFirst, data=clean)
summary(lm_perc)#not associated

lm_dose <- lm(Result24hrgL ~ F1DoseIgGfedg, data=clean)
summary(lm_dose)#not associated

lm_rep <- lm(Result24hrgL ~ Replacement, data=clean)
summary(lm_rep)#yes



#model
#tested REPLACEMENT in the model, but the model with SexMF and CrossbredYN fit the data better (lower AIC)
glmIgG24 <- lm(Result24hrgL ~ D0Fed2TxGrp  + Replacement + CrossbredYN, data=clean)
summary(glmIgG24)
plot(glmIgG24)
car::Anova(glmIgG24, type=3)


glmIgG24.2 <- lm(Result24hrgL ~ D0Fed2TxGrp  + SexMF + CrossbredYN, data=clean)
summary(glmIgG24)
plot(glmIgG24)
car::Anova(glmIgG24, type=3)

AIC(glmIgG24, glmIgG24.2)

emmeans (glmIgG24, ~ D0Fed2TxGrp)
emmeans(glmIgG24, ~CrossbredYN)
emmeans(glmIgG24, ~SexMF)
x<- emmeans (glmIgG24, ~ D0Fed2TxGrp)
pairwise_comparisons2 <- pairs(x)
summary(pairwise_comparisons2, adjust="tukey")
summary(pairwise_comparisons2)
```

## IgG 8 vs. 24hr Plot

```{r}


custom_colors <- c("MRB" = "grey30", "CSB" = "gold", "CST" = "gold3", 
                   "MCB" = "red3", "MCT" = "red4")

# Set factor orders
ordered_exp_levels <- c("Result8hrgL", "Result24hrgL")

IgG <- IgG %>%
  mutate(
    name = factor(name, levels = ordered_exp_levels),
    D0Fed2TxGrp = factor(D0Fed2TxGrp, levels = names(custom_colors))
  )

dodge_width <- 1

ggplot(IgG, aes(x = D0Fed2TxGrp, y = value, fill = D0Fed2TxGrp)) +
  theme_classic() +
  geom_boxplot(
    aes(group = interaction(D0Fed2TxGrp, name)),
    position = position_dodge(width = dodge_width),
    outlier.shape = NA
  ) +
  geom_jitter(
    aes(group = interaction(D0Fed2TxGrp, name)),
    position = position_jitterdodge(
      jitter.width = 0.5,   # horizontal spread within each dodged box
      dodge.width = dodge_width
    ),
    alpha = 0.5
  ) +
  scale_fill_manual(values = custom_colors) +
  labs(x = NULL, y = "Serum IgG (g/L)") +
  theme(
    legend.position = "none",
    text = element_text(family = "Arial", size = 12))+
  scale_y_continuous(limits = c(0, 50), breaks = seq(0, 50, by = 5))

```

## Time to Milk

```{r}
#| echo: false
#| warning: false


timetomilk <- read_csv("data/timetomilk.csv")



timetomilk |> 
  summarize (mean = mean(HrsF2toMilk), 
             sd = sd(HrsF2toMilk))

timetomilk |> 
  group_by (TRT) |> 
  summarize (mean = mean(HrsF2toMilk), 
             sd = sd(HrsF2toMilk))

timetomilk$TRT <- relevel(factor(timetomilk$TRT), ref = "MRB")
lm_timetomilk <- lm(HrsF2toMilk ~ TRT, data=timetomilk)
summary(lm_timetomilk)


```

## Change in IgG 8 to 24h

```{r}

#| echo: false
#| warning: false


clean |> 
  summarize (mean = mean(changeIgG, na.rm=TRUE), 
             sd = sd(changeIgG, na.rm=TRUE))


#one calf in the CSB group that doesn't have a 24hr blood IgG
clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(changeIgG, na.rm=TRUE), 
             sd = sd(changeIgG, na.rm=TRUE))

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_change <- lm(changeIgG ~ D0Fed2TxGrp, data=clean)
summary(lm_change)


clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = changeIgG, fill=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter(alpha = 0.5) +
  scale_fill_manual(
    values = c(
      "MRB" = "grey30",
      "CSB" = "gold",
      "CST" = "gold3",
      "MCB" = "red3",
      "MCT" = "red4")) +
  labs(x = NULL, 
       y = "Change in IgG from 8 to 24 hrs")+
  scale_y_continuous(limits = c(-20, 30), breaks = seq(-20, 30, by = 5)) +
  theme(legend.position = "none", 
        text = element_text(family = "Arial", size = 12))
 








```

### Model

```{r}
#| echo: false
#| warning: false

clean$D0Fed2TxGrp <- relevel(factor(clean$D0Fed2TxGrp), ref = "MRB")
lm_24hr <- lm(changeIgG ~ D0Fed2TxGrp, data=clean)
summary(lm_24hr) #trend 0.09

lm_24hrBreed <- lm(changeIgG ~ CrossbredYN, data=clean)
summary(lm_24hrBreed)#yes 0.07

lm_24hrSize <- lm(changeIgG ~ CalfSizeLgSm, data=clean)
summary(lm_24hrSize) #no

lm_24hrBW <- lm(changeIgG ~ BirthWtLbs, data=clean)
summary(lm_24hrBW)#no

lm_24hrRefuse <- lm(changeIgG ~ refuse, data=clean)
summary(lm_24hrRefuse)#no

lm_24hrSex <- lm(changeIgG ~ SexMF, data=clean)
summary(lm_24hrSex) #yes

lm_24hrST <- lm(changeIgG ~ SingleTwinST, data=clean)
summary(lm_24hrST)#no

lm_minold <- lm(changeIgG ~ DOFed1MinOld, data=clean)
summary(lm_minold) #no





#model
glmchange <- lm(changeIgG ~ D0Fed2TxGrp + SexMF, data=clean)
summary(glmchange)
plot(glmchange)
car::Anova(glmchange, type=3)


emmeans (glmchange, ~ D0Fed2TxGrp)
emmeans(glmchange, ~SexMF)
x<- emmeans (glmchange, ~ D0Fed2TxGrp)
pairwise_comparisons2 <- pairs(x)
summary(pairwise_comparisons2, adjust="tukey")
summary(pairwise_comparisons2)
```

## RMM Model

```{r}
#| echo: false
#| warning: false



rmm_clean$D0Fed2TxGrp <- relevel(factor(rmm_clean$D0Fed2TxGrp), ref = "MRB")

mod_base <- lme(
  fixed = IgG ~ Time * D0Fed2TxGrp + CrossbredYN + SexMF ,
  random = ~ 1 | CalfID,
  data = rmm_clean, 
  na.action = na.omit)

summary(mod_base)
car::Anova(mod_base, type=3)

emm <- emmeans(mod_base, ~ Time * D0Fed2TxGrp)
summary(emm)
pairs(emm, by = "Time")
pairs(emm, by = "D0Fed2TxGrp")

emm_df <- as.data.frame(emm)
emm_df <- emm_df %>%
  mutate(Time = factor(Time, levels = c("8hr", "24hr")))

ggplot(emm_df, aes(x = Time, y = emmean, color = D0Fed2TxGrp, group = D0Fed2TxGrp)) +
  geom_line(aes(linetype = D0Fed2TxGrp), size = 0.7) + 
  geom_point(size = 1.5) +
  scale_color_manual(
    name = NULL, 
    values = c(
      "MRB" = "grey30",
      "CSB" = "gold",
      "CST" = "gold3",
      "MCB" = "red3",
      "MCT" = "red4")) +
scale_linetype_manual(
    name = NULL,
    values = c("solid", "dotted", "dashed", "dotdash", "longdash")) + 
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  labs(
    x = "Time",
    y = "Serum IgG (g/L)") +
  theme_classic() +
  theme(legend.position = "top", 
        text = element_text(family = "Times New Roman", size = 12))+
  scale_y_continuous(limits = c(20,45), breaks = seq(20,45,5))+
  guides(color = guide_legend(override.aes = list(linetype = c("solid", "dotted", "dashed", "dotdash", "longdash"))))


```
