---
title: "Calf Health and IgG Analysis"
author: "WAK"
date: "today"
format: 
  html:
    toc: true 
    toc-float: true
    toc-location: left
---

## 

```{r}
#| include: false
#| warning: false
#| echo: false

library(epiDisplay)
library(knitr)
library(pROC)
library(tidyverse)
library(readxl)
library(ggplot2)
library(gmodels)
library(ggbeeswarm)
library(dplyr)
library(ggpubr)
library(psych)
library(lme4)
library(Matrix)
library(lmerTest)
library(irr)
library(lmerTest)
library(emmeans)
library(doBy)       
library(gmodels)    
library(car)           
library(cowplot)       
library(gridGraphics)  
library(multcomp)
library(hms)
library(gt)
```

```{r}
#| include: false
#| echo: false
#| warning: false

enroll <- read_csv('data/Zinpro 2024 enrollment final.csv')


clean <-
  enroll |> 
  filter(removeyn != 1)

#removes calf 113, who did not have correct information    

#changes these variables to "hours" vs. "seconds"
clean$DOFed2HrsOld <- as.numeric(clean$DOFed2HrsOld)/3600
clean$DOFed12IntervHrs <- as.numeric(clean$DOFed12IntervHrs)/3600
clean$DOFed1MinOld <- as.numeric(clean$DOFed1MinOld)/60
  
clean<- 
  clean |> 
  mutate(refuse=
           case_when(
             D0Fed2VolRefusemL > 0    ~ 1, 
             D0Fed2VolRefusemL == 0   ~ 0))


clean<- 
  clean |> 
  mutate(refuseM1=
           case_when(
             D1MilkRefusemL > 0    ~ 1, 
             D1MilkRefusemL == 0   ~ 0))


```

## Descriptive Analysis

### Calf Enrollment Characteristics

This generally describes calves in each treatment group by various indicators (breed, sex, size). Calves were balanced between treatment groups (n=31). Majority of enrolled calves were crossbreds, and fairly evenly split between large and small, and male and female. There was one calf who did not have their sex recorded.

```{r}
#| echo: false
#| warning: false

n_group <- 
  clean |> 
  group_by(D0Fed2TxGrp) |> 
  summarize (Count = n())


summary_sex <- 
  clean |> 
  group_by(D0Fed2TxGrp, SexMF) |> 
  summarize(Count = n())


summary_cb <- 
  clean |> 
  group_by(D0Fed2TxGrp, CrossbredYN) |> 
  summarize(Count = n())


summary_size<-
  clean |> 
  group_by(D0Fed2TxGrp, CalfSizeLgSm) |> 
  summarize(Count=n())


sick_<-
  clean |> 
  group_by(D0Fed2TxGrp, FarmDzYN) |> 
  summarize(count=n(), .groups="drop") |> 
  mutate(proportion = signif(count/sum(count), digits=2))



n_group
summary_sex
summary_cb
summary_size
sick_
```

### Birth Weight by Treatment Group

Calf birth weight was relatively consistent between treatment groups. Calves weighed between 83 and 87lbs at birth/enrollment.

```{r}
#| echo: false
#| warning: false


clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(BirthWtLbs), 
             sd = sd(BirthWtLbs))

lm_bw <- lm(BirthWtLbs ~ D0Fed2TxGrp, data=clean)
summary(lm_bw)
  

clean |> 
  group_by (D0Fed2TxGrp, CalfSizeLgSm) |> 
  summarize (mean = mean(BirthWtLbs), 
             sd = sd(BirthWtLbs))


lm_bwsize <- lm(BirthWtLbs ~ D0Fed2TxGrp*CalfSizeLgSm, data=clean)
summary(lm_bwsize)


clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = BirthWtLbs, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  labs(x = "Second Feeding Treatment", 
       y = "Birth Weight (lbs)", 
       title = "Boxplot of Treatment vs. Birth Weight")+
  scale_y_continuous(limits = c(0, 150), breaks = seq(0, 150, by = 25))
```

### Age at First Feeding(min)

```{r}
#| echo: false
#| warning: false

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(DOFed1MinOld), 
             sd = sd(DOFed1MinOld))

lm_minold <- lm(DOFed1MinOld ~ D0Fed2TxGrp, data=clean)
summary(lm_minold)

clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = DOFed1MinOld, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  labs(x = "Second Feeding Treatment", 
       y = "Age at first feeding(min)", 
       title = "Treatment vs. Age(min) at first feeding")

```

### Colostral IgG Feeding 1

```{r}
#| echo: false
#| warning: false

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(F1ColostIgGgL), 
             sd = sd(F1ColostIgGgL))

lm_F1Colost <- lm(F1ColostIgGgL ~ D0Fed2TxGrp, data=clean)
summary(lm_F1Colost)

clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = F1ColostIgGgL, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  labs(x = "Second Feeding Treatment", 
       y = "First Feeding IgG in Colostrum", 
       title = "Treatment vs. ColostIgG")+
  scale_y_continuous(limits = c(0, 125), breaks = seq(0, 125, by = 25))

```

### Feeding 1 Total IgG Dose

```{r}
#| echo: false
#| warning: false

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(F1DoseIgGfedg), 
             sd = sd(F1DoseIgGfedg))

lm_F1Dose <- lm(F1DoseIgGfedg ~ D0Fed2TxGrp, data=clean)
summary(lm_F1Dose)

clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = F1DoseIgGfedg, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  labs(x = "Second Feeding Treatment", 
       y = "First Feeding IgG Dose", 
       title = "Treatment vs. ColostIgG Dose")+
  scale_y_continuous(limits = c(0, 450), breaks = seq(0, 450, by = 50))
```

### Assn between bacteria and Colostral IgG

```{r}
#| echo: false
#| warning: false

lm_coliform <- lm(F1ColostIgGgL~F1LogColif, data=clean)
summary(lm_coliform)

clean |> 
ggplot(mapping = aes(x = F1ColostIgGgL, y = F1LogColif)) +
  theme_classic()+
  geom_point() +
  geom_smooth(method = "lm", se=TRUE, color="red")+
  labs(x = "Colostral IgG", y = "Log Coliform Count", title = "Scatterplot of Colostral IgG vs. Coliform")




lm_spc <- lm(F1ColostIgGgL~F1SPC, data=clean)
summary(lm_spc)

clean |> 
ggplot(mapping = aes(x = F1ColostIgGgL, y = F1SPC)) +
  theme_classic()+
  geom_point() +
  geom_smooth(method = "lm", se=TRUE, color="red")+
  labs(x = "Colostral IgG", y = "SPC", title = "Scatterplot of Colostral IgG vs. SPC")
```

### Brix% vs Colostral IgG

```{r}
#| echo: false
#| warning: false

lm_brix <- lm(F1ColostIgGgL~F1Brixpct, data=clean)
summary(lm_brix)

clean |> 
ggplot(mapping = aes(x = F1ColostIgGgL, y = F1Brixpct)) +
  theme_classic()+
  geom_point() +
  geom_smooth(method = "lm", se=TRUE, color="red")+
  labs(x = "Colostral IgG", y = "Brix %", title = "Scatterplot of Colostral IgG vs. Brix")
```

### Hrs Old At Second Feeding

```{r}
#| echo: false
#| warning: false

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(DOFed2HrsOld), 
             sd = sd(DOFed2HrsOld))

lm_hrsold <- lm(DOFed2HrsOld ~ D0Fed2TxGrp, data=clean)
summary(lm_hrsold)

clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = DOFed2HrsOld, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  labs(x = "Second Feeding Treatment", 
       y = "Hours old at second Feeding", 
       title = "Treatment vs. Age(hrs) at second feeding")

```

### Interval Feeding 1 - 2

```{r}
#| echo: false
#| warning: false

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(DOFed12IntervHrs), 
             sd = sd(DOFed12IntervHrs))

lm_intvhrs <- lm(DOFed12IntervHrs ~ D0Fed2TxGrp, data=clean)
summary(lm_intvhrs)

clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = DOFed12IntervHrs, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  labs(x = "Second Feeding Treatment", 
       y = "Feeding Interval", 
       title = "Treatment vs. Feeding Interval(hrs)")
```

### 8hr IgG by Treatment Group

```{r}
#| echo: false
#| warning: false

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(Result8hrgL), 
             sd = sd(Result8hrgL))

lm_8hr <- lm(Result8hrgL ~ D0Fed2TxGrp, data=clean)
summary(lm_8hr)

clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = Result8hrgL, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  labs(x = "Second Feeding Treatment", 
       y = "Serum IgG at 8 hrs", 
       title = "Treatment vs. IgG at 8 hours")+
  scale_y_continuous(limits = c(0, 60), breaks = seq(0, 60, by = 10))




```

### Refusals, second feeding (mLs and Refusal Y/N)

```{r}
#| echo: false
#| warning: false

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(D0Fed2VolRefusemL), 
             sd = sd(D0Fed2VolRefusemL))

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (refuse = sum(refuse))


lm_refuse <- lm(D0Fed2VolRefusemL ~ D0Fed2TxGrp, data=clean)
summary(lm_refuse)


clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = D0Fed2VolRefusemL, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  labs(x = "Second Feeding Treatment", 
       y = "Refusals (mLs)", 
       title = "Treatment vs. Refusals(mLs)")


```

### Second Feeding Colostral IgG

```{r}
#| echo: false
#| warning: false

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(F2ColostIgGgL), 
             sd = sd(F2ColostIgGgL))


lm_colst2 <- lm(F2ColostIgGgL ~ D0Fed2TxGrp, data=clean)
summary(lm_colst2)


clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = F2ColostIgGgL, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  labs(x = "Second Feeding Treatment", 
       y = "Second Feeding IgG", 
       title = "Treatment vs. Second Feeding IgG")
```

### Second Feeding Dose IgG

Dose this account for calves that didn't drink everything? That is why there is the weird variation?

```{r}
#| echo: false
#| warning: false

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(F2DoseIgGfedg), 
             sd = sd(F2DoseIgGfedg))


lm_colst2dose <- lm(F2DoseIgGfedg ~ D0Fed2TxGrp, data=clean)
summary(lm_colst2dose)


clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = F2DoseIgGfedg, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  labs(x = "Second Feeding Treatment", 
       y = "Second Feeding IgG dose", 
       title = "Treatment vs. Second Feeding IgG dose")
```

### Total Dose IgG Consumed (F1 + F2)

```{r}
#| echo: false
#| warning: false

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(F1F2TotIgGconsumg), 
             sd = sd(F1F2TotIgGconsumg))


lm_colstdose <- lm(F1F2TotIgGconsumg ~ D0Fed2TxGrp, data=clean)
summary(lm_colstdose)


clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = F1F2TotIgGconsumg, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  labs(x = "Second Feeding Treatment", 
       y = "Total IgG Dose", 
       title = "Treatment vs. Total IgG Dose")+
  scale_y_continuous(limits = c(0, 600), breaks = seq(0, 600, by = 100))
```

### 24hr serum IgG

```{r}
#| echo: false
#| warning: false

#one calf in the CSB group that doesn't have a 24hr blood IgG
clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(Result24hrgL, na.rm=TRUE), 
             sd = sd(Result24hrgL, na.rm=TRUE))

lm_24hr <- lm(Result24hrgL ~ D0Fed2TxGrp, data=clean)
summary(lm_24hr)


clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = Result24hrgL, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.color="black") +
  geom_jitter() +
  labs(x = "Second Feeding Treatment", 
       y = "Serum IgG at 24 hrs", 
       title = "Treatment vs. IgG at 24 hours")+
  scale_y_continuous(limits = c(0, 60), breaks = seq(0, 60, by = 10))
 


```

### First Milk (only) Feeding - Refusals

```{r}
#| echo: false
#| warning: false

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (mean = mean(D1MilkRefusemL), 
             sd = sd(D1MilkRefusemL))

clean |> 
  group_by (D0Fed2TxGrp) |> 
  summarize (refuse = sum(refuseM1))


lm_refF2 <- lm(D1MilkRefusemL ~ D0Fed2TxGrp, data=clean)
summary(lm_refF2)


clean |> 
ggplot(mapping = aes(x = D0Fed2TxGrp, y = D1MilkRefusemL, color=D0Fed2TxGrp)) +
  theme_classic()+
  geom_boxplot(outlier.shape=NA) +
  geom_jitter() +
  labs(x = "Second Feeding Treatment", 
       y = "24hr milk refusal", 
       title = "Boxplot of Treatment vs. 24hr milk refusal")
```
